What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-09 android-stacks.tex
new statistics on undocumented runtime exceptions

%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
%RSC: If needed we can remove it.
The remainder of this paper is organized as follows. Section 2 presents a
background on error handling mechanisms and Java exception handling constructs
and practices. Section 3 defines research questions, the configuration of our
exploratory study, the target applications and the evaluation procedures.
Section 4 reports our analysis of the empirical data collected relating to the
first research questions. Section X is dedicated to provide insights for the 3rd
research question. Section 6 provides further discussions and lessons learned
and discusses the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 presents our conclusions and
directions for future work.

@@

In Java an exception can be signaled in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

When an exception is signaled it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block .
The search for the handler to deal with a raised
exception occurs along the dynamic invocation chain. 
If no handler is found in the chain, the default handler associated to the 
Thread or ThreadGroup is invoked, which usually terminates the system 
and prints the exception stack trace. Figure~\ref{fig:wrapping} presents an 
example of a stack trace, on which multiple 
exception chaining is depicted. Exception Chaining (or exception wrapping) is  
common way of propagating exceptions  across components of different 
abstraction levels. It is implemented by special exception 
constructor which receives another exception as a parameter.

@@

The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.5]{stack-wrappings-hist.pdf}
  \end{center}
  \caption{Number of stack traces wrappings for Java and Android repositories. }
  \label{fig:sizewrapbhists}
\end{figure}

>>>>

%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
%RSC: If needed we can remove it.
The remainder of this paper is organized as follows. Section 2 presents a
background on exception handling mechanism. Section 3 presents the study design.
Section 4 reports study findings. Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 presents our conclusions and
directions for future work.

@@

In Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

%When an exception is signaled it causes a transfer of control from the point
%where the exception occurred to a point that can be specified by the programmer
%(exception handler) - in Java it is represented by the try-catch block .
%The search for the handler to deal with a raised
%exception occurs along the dynamic invocation chain. 
%If no handler is found in the chain, the default handler associated to the 
%Thread or ThreadGroup is invoked, which usually terminates the system 
%and prints the exception stack trace.

@@

Exception chaining is a common way of propagating exceptions in Java programs, 
which allows one exception to be wrapped in another exception and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception chaining. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.4]{stack-wrappings-hist.pdf}
  \end{center}
  \caption{Number of stack traces wrappings for Java and Android repositories. }
  \label{fig:sizewrapbhists}
\end{figure}
